---
title: "Effect of climate variables on the age-group specific malaria burden"
author: "Felix Pabon-Rodriguez"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(tinytex.verbose = TRUE)

library(dplyr) 
library(lubridate)
library(ggplot2)
library(readxl)
library(data.table)
library(tidyr)
library(lme4)
library(glmmADMB)
library(R2admb)
library(sjPlot)
library(knitr)
library(lmtest)
library(ggpubr)
library(gridExtra)
library(reshape2)
library(see)
library(MachineShop)
```

# Map of Study Sites and Clinical Facilities

\begin{center}
\includegraphics[width=\textwidth]{KenyaMap-Sites.jpg}
\end{center}

\newpage 

# Reading Malaria Positivity and Climate Variables per Sites

```{r}
# Data Site 1 (Gobei/Lwak)
s1_2022_malaria <- read_xlsx(path = "./Data per Sites/Site1_Lwak_Gobei_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 1)
s1_2023_malaria <- read_xlsx(path = "./Data per Sites/Site1_Lwak_Gobei_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 2)
s1_temp <- read_xlsx(path = "./Data per Sites/Site1_Oyude_Gobei_Temp_v2.xlsx",
                    col_names = TRUE, sheet = 1, na = c(" ", "", "*"))
s1_prec <- read_xlsx(path = "./Data per Sites/Site1-Bondo_Gobei_Precip_v2.xlsx",
                    col_names = TRUE, sheet = 1, na = c(" ", "", "*"))

# Data Site 2 (Kadengue)
s2_2022_malaria <- read_xlsx(path = "./Data per Sites/Site2-Kadenge_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 1)
s2_2023_malaria <- read_xlsx(path = "./Data per Sites/Site2-Kadenge_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 2)
s2_temp <- read_xlsx(path = "./Data per Sites/Site2_LakeAgro_Kadenge_Temp_v2.xlsx",
                   col_names = TRUE,sheet = 1, na = c(" ", "", "*"))
s2_prec <- read_xlsx(path = "./Data per Sites/Site2-Kadenge_Precip.xlsx",
                   col_names = TRUE,sheet = 1, na = c(" ", "", "*"))

# Data Site 2 (Uyawi)
s3_2022_malaria <- read_xlsx(path = "./Data per Sites/Site3-Uyawi_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 1)
s3_2023_malaria <- read_xlsx(path = "./Data per Sites/Site3-Uyawi_MalariaPositivity.xlsx",
                   col_names = TRUE,sheet = 2)
s3_temp <- read_xlsx(path = "./Data per Sites/Site3_Usenge_Uyawi_Temp_v2.xlsx",
                   col_names = TRUE,sheet = 1, na = c(" ", "", "*"))
s3_prec <- read_xlsx(path = "./Data per Sites/Site3-Uyawi_Precip.xlsx",
                   col_names = TRUE,sheet = 1, na = c(" ", "", "*"))

```

\newpage

# Data Processing (Season, Age Groups, Mean Temp, Mean Precip, Lag Values)

```{r}
## Season descriptions
seasons <- data.frame(Month = colnames(s1_2022_malaria)[-1],
                      Season = c("Dry S1", #ENE
                                 "Dry S1", #FEB
                                 "Heavy rain", #MAR
                                 "Heavy rain", #APR
                                 "Dry S2", #MAY
                                 "Dry S2", #JUN
                                 "Dry S2", #JUL
                                 "Moderate rain", #AUG
                                 "Moderate rain", #SEPT
                                 "Dry S3", #OCT
                                 "Dry S3", #NOV
                                 "Dry S3")) #DEC

# data frame to determine start and end for rectangular regions in ggplot
df <- data.frame(Season = c("Moderate rain","Dry S1", "Heavy rain", 
                            "Dry S2", "Moderate rain", "Dry S3", 
                            "Moderate rain"),
                 Start = c(0,1,3,5,8,10,12),
                 End = c(1,3,5,8,10,12,13))

##########
## Site 1
##########

# temperature and precipitation
s1_temp$Date <- as.Date(s1_temp$Date, "%m/%d/%Y")
s1_temp$month <- month(s1_temp$Date)
s1_temp$year <- year(s1_temp$Date)
 
s1_mean_temp <- s1_temp %>% 
  group_by(year,month) %>% 
  summarize(mean_temp = mean(Ave_Temperature_C ,na.rm = TRUE))
s1_mean_temp$Lag1_Temp <- lag(s1_mean_temp$mean_temp, 1)
s1_mean_temp$Lag2_Temp <- lag(s1_mean_temp$mean_temp, 2)

s1_prec$month <- month(s1_prec$Date)
colnames(s1_prec)[2:5] <- c("year2020", "year2021","year2022", "year2023")

s1_mean_prec_2021 <- s1_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2021,na.rm = TRUE))
s1_mean_prec_2021$year <- rep(2021,nrow(s1_mean_prec_2021))

s1_mean_prec_2022 <- s1_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2022,na.rm = TRUE))
s1_mean_prec_2022$year <- rep(2022,nrow(s1_mean_prec_2022))

s1_mean_prec_2023 <- s1_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2023,na.rm = TRUE))
s1_mean_prec_2023$year <- rep(2023,nrow(s1_mean_prec_2023))

s1_mean_prec <- rbind(s1_mean_prec_2021, 
                      s1_mean_prec_2022,
                      s1_mean_prec_2023)
s1_mean_prec$Lag1_Prec <- lag(s1_mean_prec$mean_prec, 1)
s1_mean_prec$Lag2_Prec <- lag(s1_mean_prec$mean_prec, 2)

s1_mean_prec <- s1_mean_prec[s1_mean_prec$year %in% c(2022,2023),
                             c("year","month","mean_prec","Lag1_Prec","Lag2_Prec")]
s1_mean_temp <- s1_mean_temp[s1_mean_temp$year %in% c(2022,2023),]

# malaria positivity
s1_d2022 <- s1_2022_malaria
s1_grouped2022 <- as.data.frame(rbind(apply(s1_d2022[1:5,2:13],2,sum),
                                   apply(s1_d2022[6:41,2:13],2,sum),
                                   apply(s1_d2022[6:15,2:13],2,sum)))
row.names(s1_grouped2022) <- c("<=5", ">5", "6-15") 

s1_d2023 <- s1_2023_malaria
s1_grouped2023 <- as.data.frame(rbind(apply(s1_d2023[1:5,2:13],2,sum),
                                   apply(s1_d2023[6:41,2:13],2,sum),
                                   apply(s1_d2023[6:15,2:13],2,sum)))
row.names(s1_grouped2023) <- c("<=5", ">5", "6-15") 

# long formatted data of malaria positivity
s1_long2022 <- as.data.frame(melt(setDT(s1_grouped2022)))
colnames(s1_long2022) <- c("Month","Count")
s1_long2022$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s1_long2022 <- merge(s1_long2022,seasons,by = "Month")
s1_long2022$Year <- rep(2022,length(s1_long2022))

s1_long2023 <- as.data.frame(melt(setDT(s1_grouped2023)))
colnames(s1_long2023) <- c("Month","Count")
s1_long2023$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s1_long2023 <- merge(s1_long2023,seasons,by = "Month")
s1_long2023$Year <- rep(2023,length(s1_long2023))

s1_malaria <- rbind(s1_long2022,s1_long2023)

##########
## Site 2
##########

# temperature and precipitation
s2_temp$Date <- as.Date(s2_temp$Date, "%m/%d/%Y")
s2_temp$month <- month(s2_temp$Date)
s2_temp$year <- year(s2_temp$Date)
 
s2_mean_temp <- s2_temp %>% 
  group_by(year,month) %>% 
  summarize(mean_temp = mean(Ave_Temperature_C,na.rm = TRUE))
s2_mean_temp$Lag1_Temp <- lag(s2_mean_temp$mean_temp, 1)
s2_mean_temp$Lag2_Temp <- lag(s2_mean_temp$mean_temp, 2)

s2_prec$month <- month(s2_prec$Date)
colnames(s2_prec)[3:5] <- c("year2021","year2022", "year2023")

s2_mean_prec_2021 <- s2_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2021,na.rm = TRUE))
s2_mean_prec_2021$year <- rep(2021,nrow(s2_mean_prec_2021))

s2_mean_prec_2022 <- s2_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2022,na.rm = TRUE))
s2_mean_prec_2022$year <- rep(2022,nrow(s2_mean_prec_2022))

s2_mean_prec_2023 <- s2_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2023,na.rm = TRUE))
s2_mean_prec_2023$year <- rep(2023,nrow(s2_mean_prec_2023))

s2_mean_prec <- rbind(s2_mean_prec_2021, 
                      s2_mean_prec_2022,
                      s2_mean_prec_2023)
s2_mean_prec$Lag1_Prec <- lag(s2_mean_prec$mean_prec, 1)
s2_mean_prec$Lag2_Prec <- lag(s2_mean_prec$mean_prec, 2)

s2_mean_prec <- s2_mean_prec[s2_mean_prec$year %in% c(2022,2023),]
s2_mean_temp <- s2_mean_temp[s2_mean_temp$year %in% c(2022,2023),]

# malaria positivity
s2_d2022 <- s2_2022_malaria
s2_grouped2022 <- as.data.frame(rbind(apply(s2_d2022[1:5,2:13],2,sum),
                                   apply(s2_d2022[6:41,2:13],2,sum),
                                   apply(s2_d2022[6:15,2:13],2,sum)))
row.names(s2_grouped2022) <- c("<=5", ">5", "6-15") 

s2_d2023 <- s2_2023_malaria
s2_grouped2023 <- as.data.frame(rbind(apply(s2_d2023[1:5,2:13],2,sum),
                                   apply(s2_d2023[6:41,2:13],2,sum),
                                   apply(s2_d2023[6:15,2:13],2,sum)))
row.names(s2_grouped2023) <- c("<=5", ">5", "6-15") 

# long formatted data of malaria positivity
s2_long2022 <- as.data.frame(melt(setDT(s2_grouped2022)))
colnames(s2_long2022) <- c("Month","Count")
s2_long2022$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s2_long2022 <- merge(s2_long2022,seasons,by = "Month")
s2_long2022$Year <- rep(2022,length(s2_long2022))

s2_long2023 <- as.data.frame(melt(setDT(s2_grouped2023)))
colnames(s2_long2023) <- c("Month","Count")
s2_long2023$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s2_long2023 <- merge(s2_long2023,seasons,by = "Month")
s2_long2023$Year <- rep(2023,length(s2_long2023))

s2_malaria <- rbind(s2_long2022,s2_long2023)


##########
## Site 3
##########

# temperature and precipitation
s3_temp$Date <- as.Date(s3_temp$Date, "%m/%d/%Y")
s3_temp$month <- month(s3_temp$Date)
s3_temp$year <- year(s3_temp$Date)
 
s3_mean_temp <- s3_temp %>% 
  group_by(year,month) %>% 
  summarize(mean_temp = mean(Ave_Temperature_C,na.rm = TRUE))
s3_mean_temp$Lag1_Temp <- lag(s3_mean_temp$mean_temp, 1)
s3_mean_temp$Lag2_Temp <- lag(s3_mean_temp$mean_temp, 2)

s3_prec$Date <- as.Date(s3_prec$Date, "%m/%d/%Y")
s3_prec$month <- month(s3_prec$Date)
colnames(s3_prec)[3:5] <- c("year2021","year2022", "year2023")

s3_mean_prec_2021 <- s3_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2021,na.rm = TRUE))
s3_mean_prec_2021$year <- rep(2021,nrow(s3_mean_prec_2021))

s3_prec$year2022 <- as.numeric(s3_prec$year2022)
s3_mean_prec_2022 <- s3_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2022,na.rm = TRUE))
s3_mean_prec_2022$year <- rep(2022,nrow(s3_mean_prec_2022))

s3_mean_prec_2023 <- s3_prec %>% 
  group_by(month) %>% 
  summarize(mean_prec = mean(year2023,na.rm = TRUE))
s3_mean_prec_2023$year <- rep(2023,nrow(s3_mean_prec_2023))

s3_mean_prec <- rbind(s3_mean_prec_2021, 
                      s3_mean_prec_2022,
                      s3_mean_prec_2023)
s3_mean_prec$Lag1_Prec <- lag(s3_mean_prec$mean_prec, 1)
s3_mean_prec$Lag2_Prec <- lag(s3_mean_prec$mean_prec, 2)

s3_mean_prec <- s3_mean_prec[s3_mean_prec$year %in% c(2022,2023),]
s3_mean_temp <- s3_mean_temp[s3_mean_temp$year %in% c(2022,2023),]

# malaria positivity
s3_d2022 <- s3_2022_malaria
s3_grouped2022 <- as.data.frame(rbind(apply(s3_d2022[1:5,2:13],2,sum),
                                   apply(s3_d2022[6:41,2:13],2,sum),
                                   apply(s3_d2022[6:15,2:13],2,sum)))
row.names(s3_grouped2022) <- c("<=5", ">5", "6-15") 

s3_d2023 <- s3_2023_malaria
s3_grouped2023 <- as.data.frame(rbind(apply(s3_d2023[1:5,2:13],2,sum),
                                   apply(s3_d2023[6:41,2:13],2,sum),
                                   apply(s3_d2023[6:15,2:13],2,sum)))
row.names(s3_grouped2023) <- c("<=5", ">5", "6-15") 

# long formatted data of malaria positivity
s3_long2022 <- as.data.frame(melt(setDT(s3_grouped2022)))
colnames(s3_long2022) <- c("Month","Count")
s3_long2022$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s3_long2022 <- merge(s3_long2022,seasons,by = "Month")
s3_long2022$Year <- rep(2022,length(s3_long2022))

s3_long2023 <- as.data.frame(melt(setDT(s3_grouped2023)))
colnames(s3_long2023) <- c("Month","Count")
s3_long2023$AgeGroup <- rep(c("<=5 years old", 
                           ">5 years old", 
                           "6-15 years old"),12)
s3_long2023 <- merge(s3_long2023,seasons,by = "Month")
s3_long2023$Year <- rep(2023,length(s3_long2023))

s3_malaria <- rbind(s3_long2022,s3_long2023)
```

\newpage

# Merge Malaria and Climate Data

```{r}
##########
## Site 1
##########

colnames(s1_mean_prec)[1:3] <- c("Year","Month","Mean_Prec")
colnames(s1_mean_temp)[1:3] <- c("Year","Month","Mean_Temp")
s1_mean_prec$Month <- rep(colnames(s1_2022_malaria)[2:13],2)
s1_mean_temp$Month <- rep(colnames(s1_2022_malaria)[2:13],2)

s1_merged_data <- merge(s1_malaria,s1_mean_temp, by=c("Year","Month"))
final_site1 <- merge(s1_merged_data,s1_mean_prec, by=c("Year","Month"))
final_site1$Site <- rep("Site1",nrow(final_site1))

##########
## Site 2
##########

s2_mean_prec <- s2_mean_prec[,c("year","month","mean_prec","Lag1_Prec",
                                "Lag2_Prec")]
colnames(s2_mean_prec)[1:3] <- c("Year","Month","Mean_Prec")

colnames(s2_mean_temp)[1:3] <- c("Year","Month","Mean_Temp")
s2_mean_prec$Month <- rep(colnames(s2_2022_malaria)[2:13],2)
s2_mean_temp$Month <- rep(colnames(s2_2022_malaria)[2:13],2)

s2_merged_data <- merge(s2_malaria,s2_mean_temp, by=c("Year","Month"))
final_site2 <- merge(s2_merged_data,s2_mean_prec, by=c("Year","Month"))
final_site2$Site <- rep("Site2",nrow(final_site2))

##########
## Site 3
##########

s3_mean_prec <- s3_mean_prec[,c("year","month","mean_prec","Lag1_Prec",
                                "Lag2_Prec")]
colnames(s3_mean_prec)[1:3] <- c("Year","Month","Mean_Prec")


colnames(s3_mean_prec)[1:3] <- c("Year","Month","Mean_Prec")
colnames(s3_mean_temp)[1:3] <- c("Year","Month","Mean_Temp")
s3_mean_prec$Month <- rep(colnames(s3_2022_malaria)[2:13],2)
s3_mean_temp$Month <- rep(colnames(s3_2022_malaria)[2:13],2)

s3_merged_data <- merge(s3_malaria,s3_mean_temp, by=c("Year","Month"))
final_site3 <- merge(s3_merged_data,s3_mean_prec, by=c("Year","Month"))
final_site3$Site <- rep("Site3",nrow(final_site3))

##########
## FINALL DATA SET
##########

analysis.data <- rbind(final_site1,
                       final_site2,
                       final_site3)
```

\newpage

# Visualizations Age-Group Malaria Burden 

```{r, fig.path='figures/'}
# visuals for malaria positivity
m1 <- ggplot(analysis.data[analysis.data$Year==2022,], aes(Month, Count , colour = AgeGroup)) +
  geom_rect(data=df, aes(NULL,NULL,xmin=Start,xmax=End,fill=Season),
            ymin=0,ymax=16000, colour="white", size=0.5, alpha=0.12) +
  scale_fill_manual(values=c("Dry S1" = "red",
                             "Dry S2" = "red",
                             "Dry S3" = "red",
                             "Moderate rain" = "green",
                             "Heavy rain" = "blue")) +
  geom_line(aes(Month, Count , group = AgeGroup), size= 1.5) +
  labs(title = "2022 Malaria Positivity Data by site",
       y = "Number of Positive Cases")+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  facet_grid(~Site) + 
  scale_colour_okabeito()
  

m2 <- ggplot(analysis.data[analysis.data$Year==2023,], aes(Month, Count , colour = AgeGroup)) +
  geom_rect(data=df, aes(NULL,NULL,xmin=Start,xmax=End,fill=Season),
            ymin=0,ymax=16000, colour="white", size=0.5, alpha=0.12) +
  scale_fill_manual(values=c("Dry S1" = "red",
                             "Dry S2" = "red",
                             "Dry S3" = "red",
                             "Moderate rain" = "green",
                             "Heavy rain" = "blue")) +
  geom_line(aes(Month, Count , group = AgeGroup), size= 1.5) +
  labs(title = "2023 Malaria Positivity Data by Site",
       y = "Number of Positive Cases")+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  facet_grid(~Site) + 
  scale_colour_okabeito()

ggarrange(m1, m2, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,
          common.legend = TRUE,
          legend = "bottom")
```

\newpage

# Visualizations of Climate Variables

```{r, fig.path='figures/'}
c1 <- ggplot(analysis.data, aes(x=Month, y=Mean_Temp, group=factor(Year))) +
  geom_line(aes(x=Month, y=Mean_Temp, color= factor(Year)),size= 1.5) +
  labs(title = "Mean Temperatures (Celsius)", color = "Year") +
  ylab("Temperature") + xlab("Month")+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  facet_grid(~Site)

c2 <- ggplot(analysis.data, aes(x=Month, y=Mean_Prec, group=factor(Year))) +
  geom_line(aes(x=Month, y=Mean_Prec, color= factor(Year)),size= 1.5) +
  labs(title = "Mean Precipitation (mm)", color = "Year") +
  ylab("Precipitation") + xlab("Month")+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  facet_grid(~Site)

ggarrange(c1, c2, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,
          common.legend = TRUE,
          legend = "bottom")

```
\newpage

# Cases per Sites and Season

```{r}
analysis.data$Month <- as.numeric(analysis.data$Month)
analysis.data$Season <- factor(analysis.data$Season, 
                               levels = c("Dry S1", 
                                          "Heavy rain", 
                                          "Dry S2", 
                                          "Moderate rain", 
                                          "Dry S3"))
analysis.data$Site <- as.factor(analysis.data$Site)

df.means <- analysis.data %>% 
  group_by(Site, AgeGroup) %>% 
  summarise(value = mean(Count))


g1 <- df.means %>% 
  ggplot() +
  aes(x = Site, y = value, color = AgeGroup) +
  geom_line(aes(group = AgeGroup),size=1.2) +
  geom_point(size=2)+
  ylab("Number of Malaria Cases")+
  ggtitle("Malaria Positivity by Sites") + 
  ylim(20,170)+ 
  scale_colour_okabeito()

df.means2 <- analysis.data %>% 
  group_by(Season, AgeGroup) %>% 
  summarise(value = mean(Count))

g2 <- df.means2 %>% 
  ggplot() +
  aes(x = Season, y = value, color = AgeGroup) +
  geom_line(aes(group = AgeGroup),size=1.2) +
  geom_point(size=2)+
  ylab("Number of Malaria Cases")+
  ggtitle("Malaria Positivity by Season") + 
  ylim(20,170) + 
  scale_colour_okabeito()


ggarrange(g1, g2, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2,
          common.legend = TRUE,
          legend = "bottom")

```

# Other Visuals

```{r}
analysis.data %>%
  group_by(Year, AgeGroup, Site) %>%
  summarise(
    Number_Cases = sum(Count)
  )

analysis.data %>%
  arrange(Site) %>%
  group_by(Year, Site) %>%
  summarise(
    Number_Cases = sum(Count)
  ) 
```


# Exploring Interactions

```{r, fig.path='figures/'}

analysis.data$Month <- as.numeric(analysis.data$Month)

mod1.data <- analysis.data[analysis.data$AgeGroup %in% c("<=5 years old",
                                                         ">5 years old"),]
mod1.data$Season <- factor(mod1.data$Season, 
                           levels = c("Dry S1", 
                                      "Heavy rain", 
                                      "Dry S2", 
                                      "Moderate rain", 
                                      "Dry S3"))
mod1.data$Site <- as.factor(mod1.data$Site)

interaction.plot(
  x.factor = mod1.data$Site,
  trace.factor = mod1.data$AgeGroup,
  response = mod1.data$Count,
  fun = median,
  ylab = "Number of Malaria Cases",
  xlab = "Sites",
  trace.label = "Age group",
  col = c("green4","blue4"),
  lyt = 1,
  lwd = 3,
  main = "Malaria Positivity by Sites")

interaction.plot(
  x.factor = mod1.data$Season,
  trace.factor = mod1.data$AgeGroup,
  response = mod1.data$Count,
  fun = median,
  ylab = "Number of Malaria Cases",
  xlab = "Season",
  trace.label = "Age group",
  col = c("green4","blue4"),
  lyt = 1,
  lwd = 3,
  main = "Malaria Positivity by Seasons")

mod2.data <- analysis.data[analysis.data$AgeGroup %in% c("<=5 years old",
                                                         "6-15 years old"),]
mod2.data$Season <- factor(mod2.data$Season, 
                           levels = c("Dry S1", 
                                      "Heavy rain", 
                                      "Dry S2", 
                                      "Moderate rain", 
                                      "Dry S3"))
mod2.data$Site <- as.factor(mod2.data$Site)

interaction.plot(
  x.factor = mod2.data$Site,
  trace.factor = mod2.data$AgeGroup,
  response = mod2.data$Count,
  fun = median,
  ylab = "Number of Malaria Cases",
  xlab = "Sites",
  trace.label = "Age group",
  col = c("green4","blue4"),
  lyt = 1,
  lwd = 3,
  main = "Malaria Positivity by Sites")

interaction.plot(
  x.factor = mod2.data$Season,
  trace.factor = mod2.data$AgeGroup,
  response = mod2.data$Count,
  fun = median,
  ylab = "Number of Malaria Cases",
  xlab = "Season",
  trace.label = "Age group",
  col = c("green4","blue4"),
  lyt = 1,
  lwd = 3,
  main = "Malaria Positivity by Seasons")
```

Comments:

- Based on interaction plots between Age group and Sites, there is a justification (as suggested in the second plot) to add an interaction effect for the comparison between <5 and 6-15 years old individuals but not for the other comparison (as shown in first plot). 

\newpage

# Exploring Goodness of Fit for Data

```{r, fig.path='figures/'}
fit.p <- glmmadmb(Count ~ AgeGroup + (1|Site),
                 data = mod1.data, 
                 zeroInflation=FALSE,
                 family="poisson", 
                 link="log")
fit.pzif <- glmmadmb(Count ~ AgeGroup + (1|Site),
                    data = mod1.data, 
                    zeroInflation=TRUE,
                    family="poisson", 
                    link="log")
fit.nb <- glmmadmb(Count ~ AgeGroup + (1|Site),
                  data = mod1.data,
                  zeroInflation=FALSE,
                  family="nbinom", 
                  link="log")
fit.zifnb <- glmmadmb(Count ~ AgeGroup + (1|Site),
                  data = mod1.data,
                  zeroInflation=TRUE,
                  family="nbinom", 
                  link="log")

anova(fit.p, fit.pzif, fit.zifnb, fit.nb)
lrtest(fit.p, fit.pzif, fit.zifnb, fit.nb)
AIC(fit.p, fit.pzif, fit.zifnb, fit.nb) 
```

Comments: 

- This analysis suggest that Negative Binomial model is a better option for modeling Malaria positivity data over time. The AIC is lower (by far) in the Negative Binomial model. 

\newpage

# Analysis of Deviance for Negative Binomial

```{r}
f1 <- glm.nb(Count ~ AgeGroup * Site, data = mod1.data)
anova(f1)
#Not significant interaction

f2 <- glm.nb(Count ~ AgeGroup * Site, data = mod2.data)
anova(f2)
#Significant interaction

f3 <- glm.nb(Count ~ AgeGroup * Season, data = mod1.data)
anova(f3)
#Not significant interaction

f4 <- glm.nb(Count ~ AgeGroup * Season, data = mod2.data)
anova(f4)
#Not significant interaction

```

\newpage

# Model Fitting

## 1. Model with age groups and no climate variables

```{r}
mod1 <- glmer.nb(Count ~ AgeGroup + Month + (1|Site) + (1|Year), 
                 data = mod1.data, 
                 control = glmerControl(optimizer = "bobyqa"))
#summary(mod1)
#exp(confint(mod1))
#lapply(ranef(mod1),exp)

mod2 <- glmer.nb(Count ~ AgeGroup + Month + (1|Site) + (1|Year), 
                 data = mod2.data, 
                 control = glmerControl(optimizer = "bobyqa"))
#summary(mod2)
#exp(confint(mod2))
#lapply(ranef(mod2),exp)
```

# Model with age groups and climate variables

```{r}
mod3 <- glmer.nb(Count ~ AgeGroup + Month + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec, 
                 data = mod1.data, 
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod3)

#data.frame(
#  estimate = exp(summary(mod3)$coefficient[,1]),
#  lower = exp(summary(mod3)$coefficient[,1] - #summary(mod3)$coefficient[,2]*1.96),
#  upper = exp(summary(mod3)$coefficient[,1] + #summary(mod3)$coefficient[,2]*1.96),
#  pval = summary(mod3)$coefficient[,4])


mod4 <- glmer.nb(Count ~ AgeGroup + Month + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec,  
                 data = mod2.data, 
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod4)

#data.frame(
#  estimate = exp(summary(mod4)$coefficient[,1]),
#  lower = exp(summary(mod4)$coefficient[,1] - #summary(mod4)$coefficient[,2]*1.96),
#  upper = exp(summary(mod4)$coefficient[,1] + #summary(mod4)$coefficient[,2]*1.96),
#  pval = summary(mod4)$coefficient[,4])
```


\newpage

## 2. A subgroup analysis focusing on 6+ (and for in 6-15)

```{r}
mod5.data <- analysis.data[analysis.data$AgeGroup == ">5 years old",]
mod5.data$Season <- factor(mod5.data$Season, 
                           levels = c("Dry S1", 
                                      "Heavy rain", 
                                      "Dry S2", 
                                      "Moderate rain", 
                                      "Dry S3"))

mod5 <- glmer.nb(Count ~ Month + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec, 
                 data = mod5.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod5)
#
#data.frame(
#  estimate = exp(summary(mod5)$coefficient[,1]),
#  lower = exp(summary(mod5)$coefficient[,1] - #summary(mod5)$coefficient[,2]*1.96),
#  upper = exp(summary(mod5)$coefficient[,1] + #summary(mod5)$coefficient[,2]*1.96),
#  pval = summary(mod5)$coefficient[,4])


mod6.data <- analysis.data[analysis.data$AgeGroup == "6-15 years old",]
mod6.data$Season <- factor(mod6.data$Season, 
                           levels = c("Dry S1", 
                                      "Heavy rain", 
                                      "Dry S2", 
                                      "Moderate rain", 
                                      "Dry S3"))

mod6 <- glmer.nb(Count ~ Month + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec, 
                 data = mod6.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod6)
#
#data.frame(
#  estimate = exp(summary(mod6)$coefficient[,1]),
#  lower = exp(summary(mod6)$coefficient[,1] - #summary(mod6)$coefficient[,2]*1.96),
#  upper = exp(summary(mod6)$coefficient[,1] + #summary(mod6)$coefficient[,2]*1.96),
#  pval = summary(mod6)$coefficient[,4])

```

\newpage

## 3. A change model where we split the year into early months and late months by using high precipitation season as the cutoff value. 

```{r}
# adding indicator for after precipitation months 
analysis.data$Post.High.Prec <- ifelse(analysis.data$Month %in% 1:4, 0, 1)
analysis.data$Post.High.Prec <- as.factor(analysis.data$Post.High.Prec)

# Subgroup analysis with high precipitation indicator
mod7b.data <- analysis.data[analysis.data$AgeGroup %in% c("<=5 years old",
                                                          ">5 years old"),]

mod7 <- glmer.nb(Count ~ AgeGroup + Month + Post.High.Prec + 
                   (1|Site) + (1|Year), 
                 data = mod7b.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod7)
#
#data.frame(
#  estimate = exp(summary(mod7)$coefficient[,1]),
#  lower = exp(summary(mod7)$coefficient[,1] - #summary(mod7)$coefficient[,2]*1.96),
#  upper = exp(summary(mod7)$coefficient[,1] + #summary(mod7)$coefficient[,2]*1.96),
#  pval = summary(mod7)$coefficient[,4])

mod8b.data <- analysis.data[analysis.data$AgeGroup %in% c("<=5 years old",
                                                          "6-15 years old"),]

mod8 <- glmer.nb(Count ~ AgeGroup + Month + Post.High.Prec + 
                   (1|Site) + (1|Year), 
                 data = mod8b.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod8)
#
#data.frame(
#  estimate = exp(summary(mod8)$coefficient[,1]),
#  lower = exp(summary(mod8)$coefficient[,1] - #summary(mod8)$coefficient[,2]*1.96),
#  upper = exp(summary(mod8)$coefficient[,1] + #summary(mod8)$coefficient[,2]*1.96),
#  pval = summary(mod8)$coefficient[,4])


# adding other climate measures
mod7.data <- analysis.data[analysis.data$AgeGroup == ">5 years old",]
mod9 <- glmer.nb(Count ~ Month + Post.High.Prec + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec, 
                 data = mod7.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod9)
#
#data.frame(
#  estimate = exp(summary(mod9)$coefficient[,1]),
#  lower = exp(summary(mod9)$coefficient[,1] - #summary(mod9)$coefficient[,2]*1.96),
#  upper = exp(summary(mod9)$coefficient[,1] + #summary(mod9)$coefficient[,2]*1.96),
#  pval = summary(mod9)$coefficient[,4])



mod8.data <- analysis.data[analysis.data$AgeGroup == "6-15 years old",]
mod10 <- glmer.nb(Count ~ Month + Post.High.Prec + (1|Site) + (1|Year) + 
                   Mean_Temp + Lag1_Temp + Lag2_Temp + 
                   Mean_Prec + Lag1_Prec + Lag2_Prec, 
                 data = mod8.data, 
                 control = glmerControl(optimizer = "Nelder_Mead",
                                        optCtrl=list(maxfun=2e5)))
#summary(mod10)
#
#data.frame(
#  estimate = exp(summary(mod10)$coefficient[,1]),
#  lower = exp(summary(mod10)$coefficient[,1] - #summary(mod10)$coefficient[,2]*1.96),
#  upper = exp(summary(mod10)$coefficient[,1] + #summary(mod10)$coefficient[,2]*1.96),
#  pval = summary(mod10)$coefficient[,4])


```


# Visualization of Results

```{r}

get_plot <- function(object = mod1){
  
  df <- data.frame(
    estimate = exp(summary(object)$coefficient[-1,1]),
    lower = exp(summary(object)$coefficient[-1,1] - summary(object)$coefficient[-1,2]*1.96),
    upper = exp(summary(object)$coefficient[-1,1] + summary(object)$coefficient[-1,2]*1.96),
    pval = summary(object)$coefficient[-1,4])
  labs <- labels(summary(object)$coefficient)[[1]][-1]
  
  p <- ggplot(df, aes(x = labs, y = estimate))+
  geom_errorbar(aes(ymax = upper, ymin = lower), width = .20)+
  geom_point()+
  geom_hline(yintercept =  1, linetype = "dotted", col = "red")+
  labs(x = "Variable", y = "Effect (95% CI)", 
       title = "Estimated Effects in the Number of Malaria Cases")+
  scale_x_discrete(labels = label_wrap_gen(10))
  
}


p1 <- get_plot(object = mod1)
p2 <- get_plot(object = mod2)
p3 <- get_plot(object = mod3)
p4 <- get_plot(object = mod4)
p5 <- get_plot(object = mod5)
p6 <- get_plot(object = mod6)
p7 <- get_plot(object = mod7)
p8 <- get_plot(object = mod8)

ggarrange(p1,p2,p3,p4,
          nrow = 2, ncol = 2, 
          common.legend = TRUE,
          labels = c("M1", "M2", "M3", "M4"))

ggarrange(p5,p6,p7,p8,
          nrow = 2, ncol = 2, 
          common.legend = TRUE,
          labels = c("M5", "M6", "M7", "M8"))


```


# Testing for Overdispersion

```{r}
test_overdispersion <- function(data){
  m <- glmer(Count ~ (1|Site) + (1|Year), 
      data = data, family = poisson,
      control = glmerControl(optimizer = "bobyqa"))
  performance::check_overdispersion(m)
}

test_overdispersion(data=mod1.data)
test_overdispersion(data=mod2.data)
test_overdispersion(data=mod5.data)
test_overdispersion(data=mod6.data)
test_overdispersion(data=mod7.data)
test_overdispersion(data=mod8.data)
test_overdispersion(data=mod7b.data)
test_overdispersion(data=mod8b.data)
```


# Model Validation / Performance

```{r}
performance::model_performance(mod1)
performance::model_performance(mod2)
performance::model_performance(mod3)
performance::model_performance(mod4)
performance::model_performance(mod5)
performance::model_performance(mod6)
performance::model_performance(mod7)
performance::model_performance(mod8)
```







